---
title: "Retention Regression Template"
output: html_document
---

#This notebook is a **template** for running the pair-specific Illinois-vs-control regression for the **Retention** analytic sample.

## What this notebook assumes

-   You are running a regression for **one adjacent year pair at a time**
-   Each row in the file already represents a **transition from year t to year t+1**
-   The dependent variable is already constructed in the analytic sample:
    -   `stayed_state_local`
-   The key treatment indicator is:
    -   `illinois` = 1 for Illinois, 0 for control states

## Interpretation

Because the file is already constructed at the transition level, this notebook estimates a **cross-sectional regression on a transition outcome** for a single year pair. This is your **pair-specific DiD-style regression** for the Retention margin.

------------------------------------------------------------------------

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(estimatr)  # for lm_robust() with HC1 SEs

# EDIT THIS when you switch to a new pair
DATA_FILE <- "analytic_sample_1011_retention.csv"

# Main variable choices
DV            <- "stayed_state_local"
TREATMENT_VAR <- "illinois"
PAIR_VAR      <- "pair"
WEIGHT_VAR    <- "weight_t"

# Baseline controls (as character strings for formula building)
BASE_CONTROLS <- c("age_t", "factor(sex_t)", "factor(race_t)", "factor(grade92_t)")

# Robust SE choice (HC1 matches statsmodels default)
COV_TYPE <- "HC1"
```

------------------------------------------------------------------------

```{r load-data, message=FALSE}
# Load the data
df <- read_csv(DATA_FILE, show_col_types = FALSE)

cat("Shape:", nrow(df), "rows x", ncol(df), "cols\n")

cat("\nPair values:\n")
print(table(df[[PAIR_VAR]], useNA = "always"))

cat("\nColumns:\n")
print(names(df))
```

------------------------------------------------------------------------

```{r validate}
# Data Validation
required_cols <- c(DV, TREATMENT_VAR, PAIR_VAR, WEIGHT_VAR)
missing <- setdiff(required_cols, names(df))

if (length(missing) > 0) {
  stop(paste("Dataset is missing required columns:", paste(missing, collapse = ", ")))
}

cat("All required columns are present.\n")
```

------------------------------------------------------------------------

```{r descriptives, message=FALSE, warning=FALSE}
# Quick descriptive checks

cat("\nUnweighted mean of DV by Illinois:\n")
df %>%
  group_by(.data[[TREATMENT_VAR]]) %>%
  summarise(mean_dv = mean(.data[[DV]], na.rm = TRUE)) %>%
  print()

cat("\nWeighted mean of DV by Illinois:\n")
df %>%
  filter(!is.na(.data[[DV]]), !is.na(.data[[WEIGHT_VAR]])) %>%
  group_by(.data[[TREATMENT_VAR]]) %>%
  summarise(
    weighted_mean = weighted.mean(.data[[DV]], w = .data[[WEIGHT_VAR]], na.rm = TRUE)
  ) %>%
  print()

if ("state_t" %in% names(df)) {
  cat("\nUnweighted mean of DV by state_t:\n")
  df %>%
    group_by(state_t) %>%
    summarise(mean_dv = mean(.data[[DV]], na.rm = TRUE)) %>%
    print()
}
```

------------------------------------------------------------------------

## Baseline specification

The default model in this notebook is a weighted Linear Probability Model:

$$Y_i = \alpha + \beta \, \text{Illinois}_i + X_i'\theta + \varepsilon_i$$

where: - $Y_i$ is `stayed_state_local` - the coefficient on `illinois` is the pair-specific Illinois-control difference for the **Retention** outcome

------------------------------------------------------------------------

```{r build-formula}
build_formula <- function(dv, controls = NULL) {
  rhs <- paste(c(TREATMENT_VAR, controls), collapse = " + ")
  as.formula(paste(dv, "~", rhs))
}

formula <- build_formula(DV, BASE_CONTROLS)
print(formula)
```

------------------------------------------------------------------------

```{r run-lpm, warning=FALSE}
run_weighted_lpm <- function(data, dv, controls = NULL,
                              weight_var = WEIGHT_VAR, cov_type = COV_TYPE) {
  controls <- controls %||% character(0)

  # Strip factor() wrappers to identify raw column names for NA dropping
  raw_controls <- gsub("^factor\\((.+)\\)$", "\\1", controls)
  needed_cols  <- unique(c(dv, TREATMENT_VAR, weight_var, raw_controls))
  needed_cols  <- intersect(needed_cols, names(data))

  before <- nrow(data)
  data   <- data %>% drop_na(all_of(needed_cols))
  after  <- nrow(data)
  cat(sprintf("Dropped %d rows due to missing values; %d rows remain.\n",
              before - after, after))

  formula <- build_formula(dv, controls)

  # lm_robust() from {estimatr} provides HC-robust SEs natively
  result <- lm_robust(
    formula  = formula,
    data     = data,
    weights  = data[[weight_var]],
    se_type  = cov_type
  )

  print(summary(result))
  result
}

lpm_result <- run_weighted_lpm(df, DV, BASE_CONTROLS)
```

------------------------------------------------------------------------

```{r result-table}
result_table <- tibble(
  term    = "illinois",
  coef    = coef(lpm_result)[["illinois"]],
  std_err = lpm_result$std.error[["illinois"]],
  p_value = lpm_result$p.value[["illinois"]],
  nobs    = lpm_result$nobs
)

result_table

install.packages("writexl")
library(writexl)
write_xlsx(result_table, "RetentionRegression1011.xlsx")


```
